noam_beta_diversity

### 2024-09-04 noam_beta_diversity.Rmd
more beta dispersion work

ggplot()+
   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
   geom_line(data = betadisper_site_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
   geom_ribbon(data = betadisper_site_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +

   labs(y = "Beta dispersion\n", x = "Year") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") )# make legend horizontal

@@

>>>>

ggplot()+
   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
   geom_line(data = betadisper_site_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
   geom_ribbon(data = betadisper_site_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +
   geom_segment(aes(y = 0.6, yend = 0.6, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   labs(y = "Beta dispersion\n", x = "Year") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") )# make legend horizontal

@@

betadisper_site_macro_mod <- glmmTMB(betadispersion_trans ~ year*habitat + (1|site), family = beta_family(link = "logit"), data = full_subset_site_macro)
summary(betadisper_site_macro_mod) # similar site variance - inv_logit(0.01) = 0.5
hist(residuals(betadisper_site_macro_mod)) # pretty good
plot(residuals(betadisper_site_macro_mod) ~ fitted(betadisper_site_macro_mod)) # slight heteroskedasticity and negative trend
Anova(betadisper_site_macro_mod)
#               Chisq Df Pr(>Chisq)    
# year         36.591  1  1.457e-09 ***
# habitat      46.029  3  5.590e-10 ***
# year:habitat 31.232  3  7.597e-07 ***

emtrends(betadisper_site_macro_mod, pairwise ~ habitat , "year") # backreef distinct from all other habitats
betadisper_site_macro_emm <- emmip(betadisper_site_macro_mod, habitat ~ year , at = list(year = seq(2007,2023,1)), plotit = F, CIs = T)

ggplot()+
   geom_point(data = site_full, aes(x = year, y = betadispersion_trans, colour = habitat), alpha = 0.1 ) +
   geom_line(data = betadisper_site_macro_emm, aes(x = year, y = inv_logit(yvar), colour = habitat), linewidth = 2) +
   geom_ribbon(data = betadisper_site_macro_emm, aes(x = year, ymin = inv_logit(LCL), ymax = inv_logit(UCL), fill = habitat), alpha = 0.3) +
   scale_colour_manual(values = habitat_colours) +
   scale_fill_manual(values = habitat_colours) +
   geom_segment(aes(y = 0.75, yend = 0.75, x = 2006, xend = 2010), linewidth = 2, alpha = 1, colour = "grey") +
   geom_vline(xintercept = 2010, linetype = 2, colour = "black", size = 1) +
   geom_vline(xintercept = 2019, linetype = 2, colour = "red", size = 1) +
   labs(y = "Beta dispersion\n", x = "Year") +
   theme_bw() + 
   theme(panel.grid.major = element_blank(), 
         panel.grid.minor = element_blank(),
         panel.background = element_blank(),
         plot.title = element_text(color = "grey20", size = 32, vjust = 0, face = "plain"),
         axis.line.y = element_line(colour = "black"),
         axis.text.y = element_text(color = "grey20", size = 28, hjust = .5, vjust = .5, face = "plain"),
         axis.title.y = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         axis.line.x = element_line(colour = "black"),
         axis.text.x = element_text(color = "grey20", size = 28, angle = 90, vjust = 0.5, hjust=1, face = "plain"),
         axis.title.x = element_text(color = "grey20", size = 32, hjust = .5, vjust = 0, face = "plain"),
         strip.text.x = element_text(size = 32),
         legend.key.size = unit(2, "cm"), # increase legend size
         legend.text = element_text(size = rel(2)), # increase legend text
         legend.title = element_text(size = rel(2.25)),
         legend.position = 'bottom', # move legend to bottom
         legend.box = 'horizontal',
         legend.box.background = element_rect(colour = "black") )# make legend horizontal
